TacticToe: Learning to Prove with Tactics

### 2018-08-20 tactictoe_jv.tex
Term instantiation

Here, the two examples accepts only very specific kind of theorems: the theorem 
argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
and an induction principle should be given to \texttt{INDUCT\_THEN}.
The likelihood of finding such theorem is low, thus given the list of 
predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
[thm1,\ldots,thm16]$, we create 16 instantiated tactics and try them in the 
order given by the predictor, and we return the result of the first successful 
tactic and produces new goals. For readers familiar with \holfour tacticals, we 
give the single new tactic that can be created using this procedure on the 
first example:  
  \texttt{CHANGED\_TAC (MATCH\_MP\_TAC thm1) ORELSE ... ORELSE CHANGED\_TAC 
  (MATCH\_MP\_TAC thm16)}.

\paragraph{Term}
We re-use the scheme describe previously for changing the prediction of one 
argument into the prediction of a list of arguments.
The only things that changes is that we have to design a special predictor for 
terms. In this context we think that the predictions should come from the list 
of subterms present in the goal as a first approximation. So all algorithm will 
try

Side experiments with term argument predictions 
show that trying to instantiate the tactic with subterms 
from the goal improve the success rate by another 4 percent. 
However, in the evaluation presented in this paper, the 
abstraction/instantiation method is evaluated only for the type of theorem list.
Moreover, the current argument predictions are only influenced by the input 
goal, 
and we believe that in the future it would be beneficial to have more flexible 
predictors that can adapt their predictions to each abstracted tactic.

>>>>

Here, the two examples accepts only very specific kind of theorems: the theorem 
argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
and an induction principle should be given to \texttt{INDUCT\_THEN}.
The likelihood of finding such theorem is low, thus given the list of 
predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
[thm1,\ldots,thm16]$, we create 16 instantiations and try them in the 
order given by the predictor, and we return the result of the first successful 
tactic and produces new goals. With a modified version of 
the tactical \texttt{ORELSE} that fails if the tactic 
did not have any effect, we
can write for the first example a single new tactic encoding this procedure: 
  \[\texttt{MATCH\_MP\_TAC thm1 ORELSE ... ORELSE MATCH\_MP\_TAC thm16}\]

\paragraph{Term}
Some proofs requires to instantiate a tactic argument by a term. Here are two 
abstracted 
tactics that uses term 
arguments with placeholders \texttt{term} and \texttt{thm}: 
  \[\texttt{EXISTS\_TAC term},\ \texttt{SPEC\_THEN term STRUCT\_CASES\_TAC 
  thm}\]
Finding a witness to instantiate an existential formula is difficult because it 
is likely that this witness may not have been proposed before and there is no 
general rule on how to generate it. Instantiating 
the term argument in the second tactic is easier because the expected term is a 
variable of the goal. That is why we think that the first step toward term 
prediction is to select subterms of the goal ordered by their similarity with 
the term argument appearing in the original tactic. 
Side experiments using a tactic that tries the 16 closest subterms 
$[term1,\ldots,term16]$ to the goal with the procedure described in the 
previous paragraph 
shows that this simple term predictor improves the success rate of 
\tactictoe by another 4 percent. The tactic derived from the abstraction
\texttt{EXISTS\_TAC term} in this experiment is:
   \[\texttt{EXISTS\_TAC term1 ORELSE ... ORELSE EXISTS\_TAC term16}\]

In the future, with stronger machine learning models, we should be able to 
select the right term from larger sets of terms including:
\begin{itemize}
 \item subterms of the goal (already implemented),
 \item terms appearing as arguments of the abstracted tactic,
 \item tactic sub-expressions of type term and their subterms,
 \item subterms of theorems from the database $\mathbb{D}_{\mathit{theorem}}$,
 \item terms created by term synthesis techniques.
\end{itemize}

\begin{remark} (Limitations)
In the experiments presented in Section~\ref{sec:evaluation}, the 
abstraction/instantiation method is evaluated only for the type ``theorem 
list''.
Another issue is that the argument predictions are only influenced by the input 
goal. But we believe that it would be beneficial to have
more flexible predictors that can adapt their selection to the abstracted 
tactic they procure instantiations for.
\end{remark}
